On probabilistic one-vs-one classification

### 2025-04-21 paper.tex
Fix bibliography

\usepackage{multirow}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{tikz-cd}

@@

A key step in probabilistic one-vs-one modelling is the aggregation of results of individual binary classifiers - so-called \emph{coupling method}. The most commonly used coupling method is that proposed by Wu-Lin-Weng. As noted in \cite{dohau}, the method is one of several non-canonical decisions used in multi-class SVM modelling. The goal of this paper is to compare its behavior with canonical methods.

@@

\emph{Bayes covariance} is a notion introduced in  \cite{vsuch2016bayes}. It refers to the behavior of a probabilistic multi-class method built in one-vs-one fashion, when the priors change. Any Bayes classifier, whether binary or multiclass changes its prediction when the priors change. Let us describe the process. Suppose we change priors from $\Pi = (P_1, \ldots, P_k)$ to $\Pi'= (P'_1, \ldots, P'_K)$. Then the prediction of Bayes classifiers changes from $\boldsymbol{p}^\Pi= (p_1, \ldots, p_K)$ to the vector 
\begin{align}
\boldsymbol{p}^{\Pi'} \propto (p_1 \frac{P'_1}{P_1}, \ldots, p_K \frac{P'_K}{P_K}). \label{eq:changePrior}
\end{align}

@@

Let us finally mention \emph{Hinton's minimality}. Geoffrey Hinton is reported to object to the underlying principle of one-vs-one classification, where any single classifier influences the multi-class prediction \cite[p.~467]{hastie1998classification}. Suppose we are classifying objects from CIFAR-10 dataset.  If the true class is  a ``dog'', why should we care what is the output of the classifier distinguishing ``airplanes'' from ``ships''? The binary classifier is trained only on airplanes and ships, and has never seen a dog. One may thus prefer a classifier which somehow suppresses noisy output of (seemingly) irrelevant classifiers.

@@

A notable example which uses this principle is the coupling method introduced in \cite{hastie1998classification}. Their method uses Kullback-Leibler divergence $\divm_\textrm{KL}$
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{p}, \boldsymbol{q})= \sum p_i \log (p_i/ q_i)
\end{align*}

@@

As an example, consider the classification problem with three classes. We have three arboreal classifiers $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ which correspond to star graphs centered respectively at classes $C_1, C_2, C_3$. The classifier 
\begin{align}
\frac13 \boldsymbol{s}_1 \oplus \frac13 \boldsymbol{s}_2 \oplus \frac 13 \boldsymbol{s}_3  \label{eq:bc1}
\end{align}
is Bayes covariant \emph{and} symmetric. A key result proved in \cite{vsuch2016bayes} is that it is a unique such classifier for $K=3$.

@@

It is natural to consider the benchmark datasets used in previous evaluation of coupling methods \cite{wu2004probability}. The authors considered altogether seven datasets with number of classes ranging from 3 to 26. For each dataset they extracted 40 pairs of training and testing datasets. Half of these were smaller (300 training samples and 500 testing samples) and the other half was larger (800 training samples and 1000 testing samples). In our experiments we use only the larger pairs.

@@

We have opted to use LDA as a binary classification tool. Its appeal is that it is applicable even for linearly separable datasets, thus obviating additional crossvalidation step incurred by methods that use regularization (e.g. penalized logistic regression or SVM). It usually provides results close to logistic regression \cite{james2013introduction}.

@@

Although previous evaluations of the latter two methods were done in \cite{vsuch2016bayes}, they used phonetic data for which ground truth may be imprecise.

@@

In the third step we define two parameter family of Bayes covariant coupling methods
\begin{align}
a_1 \boldsymbol{s}_1 + a_2 \boldsymbol{s}_2 + a_3 \boldsymbol{s}_3,\quad\textrm{where } a_1 + a_2 +a_3 = 1
\end{align}
Note that the parameter space of this family is not compact. In order to facilitate computations we opted to conduct a grid search on points in the simplex where the coefficients $a_i$ are nonnegative $a_i \geq 0$.

In order to reduce computational burden we selected 1000 three-class subproblems. The results are shown in Figure \ref{fig:par-bc}.

@@

\begin{figure}[!ht]
	\includegraphics{graph/exp2-detail.pdf}
	\caption{Color map indicates the value of Brier score computed for an ensemble of coupling methods on the testing dataset. The vertices triangle indicate the arboreal coupling methods, and the black circle the location of the optimum.}
\end{figure}

>>>>

\usepackage{multirow}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{tikz-cd}
\usepackage[utf8]{inputenc} % Enable UTF-8 encoding
\usepackage[T1]{fontenc}
\usepackage{textgreek}

@@

A key step in probabilistic one-vs-one modelling is the aggregation of results of individual binary classifiers - so-called \emph{coupling method}. The most commonly used coupling method is that proposed by Wu-Lin-Weng. As noted in [\cite{dohau}], the method is one of several non-canonical decisions used in multi-class SVM modelling. The goal of this paper is to compare its behavior with canonical methods.

@@

\emph{Bayes covariance} is a notion introduced in  [\cite{vsuch2016bayes}]. It refers to the behavior of a probabilistic multi-class method built in one-vs-one fashion, when the priors change. Any Bayes classifier, whether binary or multiclass changes its prediction when the priors change. Let us describe the process. Suppose we change priors from $\Pi = (P_1, \ldots, P_k)$ to $\Pi'= (P'_1, \ldots, P'_K)$. Then the prediction of Bayes classifiers changes from $\boldsymbol{p}^\Pi= (p_1, \ldots, p_K)$ to the vector 
\begin{align}
\boldsymbol{p}^{\Pi'} \propto (p_1 \frac{P'_1}{P_1}, \ldots, p_K \frac{P'_K}{P_K}). \label{eq:changePrior}
\end{align}

@@

Let us finally mention \emph{Hinton's minimality}. Geoffrey Hinton is reported to object to the underlying principle of one-vs-one classification, where any single classifier influences the multi-class prediction [\cite[p.~467]{hastie1998classification}]. Suppose we are classifying objects from CIFAR-10 dataset.  If the true class is  a ``dog'', why should we care what is the output of the classifier distinguishing ``airplanes'' from ``ships''? The binary classifier is trained only on airplanes and ships, and has never seen a dog. One may thus prefer a classifier which somehow suppresses noisy output of (seemingly) irrelevant classifiers.

@@

A notable example which uses this principle is the coupling method introduced in [\cite{hastie1998classification}]. Their method uses Kullback-Leibler divergence $\divm_\textrm{KL}$
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{p}, \boldsymbol{q})= \sum p_i \log (p_i/ q_i)
\end{align*}

@@

As an example, consider the classification problem with three classes. We have three arboreal classifiers $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ which correspond to star graphs centered respectively at classes $C_1, C_2, C_3$. The classifier 
\begin{align}
\frac13 \boldsymbol{s}_1 \oplus \frac13 \boldsymbol{s}_2 \oplus \frac 13 \boldsymbol{s}_3  \label{eq:bc1}
\end{align}
is Bayes covariant \emph{and} symmetric. A key result proved in [\cite{vsuch2016bayes}] is that it is a unique such classifier for $K=3$.

@@

It is natural to consider the benchmark datasets used in previous evaluation of coupling methods [\cite{wu2004probability}]. The authors considered altogether seven datasets with number of classes ranging from 3 to 26. For each dataset they extracted 40 pairs of training and testing datasets. Half of these were smaller (300 training samples and 500 testing samples) and the other half was larger (800 training samples and 1000 testing samples). In our experiments we use only the larger pairs.

@@

We have opted to use LDA as a binary classification tool. Its appeal is that it is applicable even for linearly separable datasets, thus obviating additional crossvalidation step incurred by methods that use regularization (e.g. penalized logistic regression or SVM). It usually provides results close to logistic regression [\cite{james2013introduction}].

@@

Although previous evaluations of the latter two methods were done in [\cite{vsuch2016bayes}], they used phonetic data for which ground truth may be imprecise.

@@

In the third step we define two parameter family of Bayes covariant coupling methods
\begin{align}
a_1 \boldsymbol{s}_1 \oplus a_2 \boldsymbol{s}_2 \oplus a_3 \boldsymbol{s}_3,\quad\textrm{where } a_1 + a_2 +a_3 = 1 \label{eq:family}
\end{align}
Note that the parameter space of this family is not compact. In order to facilitate computations we opted to conduct a grid search on points in the simplex where the coefficients $a_i$ are nonnegative $a_i \geq 0$.  To further reduce computational burden we selected 1000 three-class subproblems from those where Wu-Lin-Weng's method outperformed normal coupling. The results are shown in Figure \ref{fig:par-bc}.

@@

We can see that only in minority of cases we found a parametric method matching the performance of Wu-Lin-Weng's. This finding comes with caveat that we searched only in the simplex $a_i \geq 0$.

Therefore as the final step in our investigation we looked whether a parametric Bayes covariant method in family \eqref{eq:family} for parameters \emph{outside} of this simplex could perform better. We conducted another search but this time
\begin{itemize}
	\item We searched within a much larger rectangle $D$ containing the simplex $a_i\geq 0$, which in this case is a triangle $T$.
	\item We used Brier score as a smooth proxy for measuring the quality of multi-class classifier. This score is much smoother than 0-1 score which measures accuracy.
\end{itemize}

The results are shown in Figure \ref{fig:score}

\begin{figure}[!ht]
	\includegraphics{graph/exp2-detail.pdf}
	\caption{Color map indicates the value of Brier score computed for an ensemble of coupling methods on the testing dataset. The vertices triangle indicate the arboreal coupling methods, and the black circle the location of the optimum.}
	\label{fig:score}
\end{figure}

We can see that indeed the optimum over the rectangle lies outside of the simplex. Let us compare the resulting accuracy of the optima in $\Delta$ and $D$. The results are shown in Table \ref{tab:step5}.

\input tab-step5.tex
